---
title: "Chi-Square Test for Daily Bar Sales"
format:
  html:
    code-fold: true
---

```{r}
library(tidyverse)
library(vcd)
```

### Load the Data:

```{r}
bar_sales <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/main/bar_sales.csv")
```


## Scenario

You received a tip that a local bar is cheating the tax collector. After interviewing the bar owner, you learn that the bar's weekly sales total $49,000, and the owner claims that the daily traffic for the week is broken down into the following percentages, on which he pays sales taxes on his weekly total:

```{r}
bar_sales |> 
  select(day, percentage, expected)
```

However, you obtained the actual daily sales observations for the week from that disgruntled accountant:

```{r}
bar_sales |> 
  select(day, observed_percentage, observed_sales)
```


### Chi-Square Test

To determine if the observed daily sales align with the expected percentages reported by the bar owner, we will perform a chi-square test:

```{r}
chisq_result <- chisq.test(bar_sales$observed_sales, p = bar_sales$percentage)
chisq_result
```

```{r}
alpha <- 0.10
df <- length(bar_sales$day) - 1
critical_value <- qchisq(alpha, df, lower.tail = FALSE)
critical_value
```

