```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

## Identifying Schools with Biggest Differences in Math Test Scores

### Introduction

In this analysis, we will use linear regression to identify the schools that show the biggest differences between their actual and expected math test scores. We'll then create a scatterplot to visualize the relationship between the FARMS (Free and Reduced-price Meal System) percentage and the percentage of students who met or exceeded math expectations.

This analysis can be useful for journalists and policymakers to understand which schools are performing better or worse than expected, based on the socioeconomic status of their student population. By identifying these schools, targeted interventions and support can be provided to help improve student outcomes.

### Loading and Preparing the Data

First, let's load the data into a data frame using the read_csv() function from the tidyverse library.

```{r}
schools <- read_csv("montgomery_pa.csv")
```

### Performing Linear Regression

Next, we'll perform a linear regression analysis using the lm() function. We'll set the percentage of students who met or exceeded math expectations as the dependent variable, and the FARMS percentage as the independent variable.

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)

summary(model)
```

### Identifying Schools with Biggest Residuals

To find the schools with the biggest differences between their actual and expected math test scores, we'll extract the residuals from the linear regression model. The residuals represent the difference between the actual and predicted values.

We'll then use the arrange() function to sort the data frame by the residuals, both in descending and ascending order, and take the top 5 schools in each direction.

```{r}
schools$predicted <- predict(model)

# Calculate the residuals
schools$residual <- schools$`% Met or Exceeded Math Expectations` - schools$predicted

# Sort the data frame by the absolute value of the residuals in descending order
schools <- schools %>%
  mutate(abs_residual = abs(residual)) %>%
  arrange(desc(abs_residual))

```
The schools with the biggest positive residuals are performing better than expected, while the schools with the biggest negative residuals are performing worse than expected, based on their FARMS percentage.

### Creating a Scatterplot

Finally, let's create a scatterplot to visualize the relationship between the FARMS percentage and the percentage of students who met or exceeded math expectations. We'll use ggplot2 to create the plot and add a linear regression line.

```{r}
schools %>%
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations", title = "Scatterplot of Math Expectations vs FARMS Percentage")
```

The scatterplot shows the overall trend of the relationship between the FARMS percentage and the math test scores. The red line represents the linear regression line, which helps us understand the average relationship between the two variables.

By identifying the schools with the biggest residuals, we can further investigate the factors that may be contributing to their performance, either positive or negative, compared to the general trend.

## Describe the Plot

What's the general trend described by this model and the scatterplot? Write it in the form of a sentence that describes what happens when one of the variables increases.
