```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Testing Statistical Significance with P-values

### Are Maryland's WBB players shorter than other D-I teams' players?

```{r}
# Load the data

wbb_rosters <- read_csv("https://raw.githubusercontent.com/Sports-Roster-Data/womens-college-basketball/main/wbb_rosters_2023_24.csv") |> filter(!is.na(total_inches)) |> filter(division == 'I')

sample_data <- wbb_rosters[sample(nrow(wbb_rosters), 100), ]

maryland <- wbb_rosters |> filter(ncaa_id == 392)
```

#### Define hypotheses

Null Hypothesis (H0): There is no significant difference in the average height of players on Maryland compared to the average height of a sample of players from the NCAA's Division I.

Alternative Hypothesis (H1): There is a significant difference in the average height of players on Maryland compared to the average height of a sample of players from the NCAA's Division I.


```{r}

# Set significance level
alpha <- 0.05

# Perform the hypothesis test
result <- t.test(maryland$total_inches, mu = mean(sample_data$total_inches))

# Calculate p-value
p_value <- result$p.value

# Make a decision
if (p_value < alpha) {
  decision <- "Reject the null hypothesis"
} else {
  decision <- "Fail to reject the null hypothesis"
}

# Interpretation
interpretation <- paste("Based on the p-value of", round(p_value, 4),
                        ", at a significance level of", alpha,
                        ", we", decision, "and conclude that there is",
                        "a significant difference in the average height of players",
                        "on Maryland compared to the average height of players across",
                        "all teams in the league.")

interpretation

```

### Prince George's County Crime Data

```{r}
pg_crimes <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/main/Crime_Incidents_July_2023_to_Present_20240414.csv") |> clean_names()

pg_crimes <- pg_crimes |> 
  mutate(clean_date = mdy_hms(date), month = month(clean_date), week_day = wday(clean_date))

pg_crimes |> 
  filter(str_detect(clearance_code_inc_type, "AUTO, STOLEN")) |> 
  group_by(month) |> 
  summarize(total = n())
```

### Questions about crimes over time: 

#### Stolen cars by month: was March's decline statistically significant?

Null Hypothesis: 
Alternative Hypothesis: 

### Set up our t.test by creating an average based on pre-March figures

```{r}

car_thefts <- pg_crimes |> 
  filter(!month %in% c(3, 4)) |>  # April isn't finished yet
  filter(str_detect(clearance_code_inc_type, "AUTO, STOLEN")) |> 
  group_by(month) |> 
  summarize(total = n())

# Set significance level
alpha <- 0.05

# Get summary stats
car_thefts |> 
  summarize(mean = mean(total), sd = sd(total))

# Perform the hypothesis test
result <- t.test(car_thefts$total, mu = 358)

# Calculate p-value
p_value <- result$p.value

# Make a decision
if (p_value < alpha) {
  decision <- "Reject the null hypothesis"
} else {
  decision <- "Fail to reject the null hypothesis"
}

# Interpretation
interpretation <- paste("Based on the p-value of", round(p_value, 4),
                        ", at a significance level of", alpha,
                        ", we", decision)

interpretation
result
```

#### What about murders?

Null Hypothesis: There has been no significant change in murders this year.
Alternative Hypothesis: There has been a significant change in murders this year.

### Set up our data

```{r}
# Given data
homicides <- pg_crimes |> 
  filter(month != 4) |>  # April isn't finished yet
  filter(clearance_code_inc_type == 'HOMICIDE') |> 
  group_by(month) |> 
  summarize(total = n())

# Set significance level
alpha <- 0.05

# Calculate the sample mean
homicides |> 
  summarize(mean = mean(total), sd = sd(total))

# Perform one-sample t-test
test_result <- t.test(homicides, mu = 7)

# Extract p-value
p_value <- test_result$p.value

# Make decision based on p-value
if (p_value < alpha) {
  decision <- "reject the null hypothesis"
} else {
  decision <- "fail to reject the null hypothesis"
}

# Print the results
cat("Results of the significance test for change in murders by month:\n")
cat("Test statistic:", round(test_result$statistic, 2), "\n")
cat("p-value:", round(p_value, 4), "\n")
cat("Conclusion:", decision, "at the", alpha, "significance level.\n")


```


